### 2021-06-03 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Fixed spacing and margins for eqns in ch4. Need to do same for ch3.

		& \quad \quad + \alphaa_{n,k,3} \:  \scopa_{n, k-1, i}(x, y, z) + \alphaa_{n,k,4} \:  \scopa_{n, k+1, i}(x, y, z) \nonumber \\
		& \quad \quad + \alphaa_{n,k,5} \:  \scopa_{n+1, k-1, i}(x, y, z) + \alphaa_{n,k,6} \:  \scopa_{n+1, k+1, i}(x, y, z), \\ \\

@@

		& \quad \quad + \betaa_{n,k,i,3} \:  \scopa_{n, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,4} \:  \scopa_{n, k+1, |i-1|}(x, y, z) \nonumber \\
		& \quad \quad + \betaa_{n,k,i,5} \:  \scopa_{n+1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,6} \:  \scopa_{n+1, k+1, |i-1|}(x, y, z), \\ \\
	z \: \scopnkia(x,y,z) &= \gammaa_{n,k,1} \: \scopa_{n-1, k, i}(x, y, z) + \gammaa_{n,k,2} \: \scopa_{n, k, i}(x, y, z) + \gammaa_{n,k,3} \: \scopa_{n+1, k, i}(x, y, z),

@@

Recall from \bsrefeqn{eqn:sc:normscop} that $\norm{\scopmjha}_{\scopa}^2 = \pi \: \normgenjac^{(a,2j)}$. Then for $m = 0,\dots,n+1$, $j = 0,\dots,m$, using a change of variables $(\cos\theta \sin\varphi, \: \sin\theta\sin\varphi, \: \cos\varphi) = (x, y, z)$:

@@

	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta \Big) \\

@@

	&= \half \: \pi \: \delta_{i,h} \: (\eta_{k-1} \: \delta_{k-1, j} + \eta_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.

@@

	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \cdot \: \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta \Big) \\

@@

	&= \half \: \pi \: \delta_{|i-1|,h} \: \big[(-1)^{i+1} \: \eta_{k-1} \: \delta_{k-1, j} + (-1)^i \: \eta_{k} \: \delta_{k+1, j}\big] \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z.

@@

Each of these Jacobi matrices are then block-tridiagonal (block-bandwidths $(1,1)$). For $J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
	A^{(a)}_{x,n} &:= 
		\begin{pmatrix}
			0 & A^{(a)}_{n,0,6} & 0 & & \\
			A^{(a)}_{n,1,5} & \ddots & \ddots & & \\
			& \ddots & \ddots & \ddots & \\
			& & A^{(a)}_{n,n,5} & 0 & A^{(a)}_{n,n,6} \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
	B^{(a)}_{x,n} &:= 
		\begin{pmatrix}
			0 & A^{(a)}_{n,0,4} & & \\
			A^{(a)}_{n,1,3} & \ddots & \ddots & \\
			& \ddots & \ddots & A^{(a)}_{n,n-1,4} \\
			& & A^{(a)}_{n,n,3} & 0
		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
	C^{(a)}_{x,n} &:= 
		\begin{pmatrix}
			0 & A^{(a)}_{n,0,2} & & \\
			A^{(a)}_{n,1,1} & \ddots & \ddots & \\
			& \ddots & \ddots &A^{(a)}_{n,n-2,2} \\
			& & \ddots & 0 \\
			& & & A^{(a)}_{n,n,1} \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
where for $k = 1,\dots,N, \: n = k,\dots,N$
	A^{(a)}_{n,k,j} &:= 
		\begin{pmatrix}
			\alphaa_{n,k,j} & 0 \\
			0 & \alphaa_{n,k,j}
		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
	A^{(a)}_{n,0,j} &:=
		\begin{pmatrix}
			\alphaa_{n,0,j} & 0
		\end{pmatrix} \in \R^{1\times2}, j \text{ even} \\
	A^{(a)}_{n,1,j} &:=
		\begin{pmatrix}
			\alphaa_{n,1,j} \\
			0
		\end{pmatrix} \in \R^{2\times1}, j \text{ odd}.
For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
	A^{(a)}_{y,n} &:= 
		\begin{pmatrix}
			0 & B^{(a)}_{n,0,6} & 0 & & \\
			B^{(a)}_{n,1,5} & \ddots & \ddots & & \\
			& \ddots & \ddots & \ddots & \\
			& & B^{(a)}_{n,n,5} & 0 & B^{(a)}_{n,n,6} \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
	B^{(a)}_{y,n} &:= 
		\begin{pmatrix}
			0 & B^{(a)}_{n,0,4} & & \\
			B^{(a)}_{n,1,3} & \ddots & \ddots & \\
			& \ddots & \ddots & B^{(a)}_{n,n-1,4} \\
			& & B^{(a)}_{n,n,3} & 0
		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
	C^{(a)}_{y,n} &:= 
		\begin{pmatrix}
			0 & B^{(a)}_{n,0,2} & & \\
			B^{(a)}_{n,1,1} & \ddots & \ddots & \\
			& \ddots & \ddots & B^{(a)}_{n,n-2,2} \\
			& & \ddots & 0 \\
			& & & B^{(a)}_{n,n,1} \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
where for $k = 1,\dots,N, \: n = k,\dots,N$  
	B^{(a)}_{n,k,j} &:= 
		\begin{pmatrix}
			0 & \betaa_{n,k,0,j} \\
			\betaa_{n,k,1,j} & 0
		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
	B^{(a)}_{n,0,j} &:=
		\begin{pmatrix}
			0 & \betaa_{n,0,0,j}
		\end{pmatrix} \in \R^{1\times2}, j \text{ even} \\
	B^{(a)}_{n,1,j} &:=
		\begin{pmatrix}
			0 \\
			\betaa_{n,1,1,j}
		\end{pmatrix}\in \R^{2\times1}, j \text{ odd}.
For $J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
	A^{(a)}_{z,n} &:= 
		\begin{pmatrix}
			\Gamma^{(a)}_{n,0,3} & 0 & \\
			0 & \ddots & \ddots & & \\
			& \ddots & \ddots & \ddots & \\
			& & 0 & \Gamma^{(a)}_{n,n,3} & 0 \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)}, \quad n = 0,1,2,\dots \\
	B^{(a)}_{z,n} &:= 
		\begin{pmatrix}
			\Gamma^{(a)}_{n,0,2} & \\
			& \ddots & & \\
			& & \ddots & \\
			& & & \Gamma^{(a)}_{n,n,2}
		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)}  \quad n = 0,1,2,\dots \\
	C^{(a)}_{z,n} &:= 
		\begin{pmatrix}
			\Gamma^{(a)}_{n,0,1} & 0 & & \\
			0 & \ddots & \ddots & \\
			& \ddots & \ddots & 0 \\
			& & \ddots & \Gamma^{(a)}_{n,n-1,1} \\
			& & & 0 \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad n = 1,2,\dots
where for $k = 1,\dots,N, \: n = k,\dots,N$
	\Gamma^{(a)}_{n,k,j} &:= 
		\begin{pmatrix}
			\gamma_{n,k,j} & 0 \\
			0 & \gamma_{n,k,j}
		\end{pmatrix} \in \R^{2\times2}, \label{eqn:sc:jacobisubblocksGamma1} \\
	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}. \label{eqn:sc:jacobisubblocksGamma2}
Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.

@@

A multivariate analogue of Clenshaw's algorithm for recursive evaluation of the OPs has been established for methods on other domains in 2D such as the triangle \cite{olver2019triangle} and the disk-slice \cite{snowball2019sparse}, and we proceed similarly here. Combining each system in \bsrefeqn{eqn:sc:jacobimatricesdefinitionalt} into a block-tridiagonal system, for any $(x,y,z) \in \Omega$, yields:
\bseqn{
	\renewcommand\arraystretch{1.3}
	\begin{pmatrix}
		1 & & & \\
		B_0-G_0(x,y,z) & A_0 & & \\
		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
		& C_2 & B_2 - G_2(x,y,z)  & \ddots \\
		& & \ddots &\ddots
	\end{pmatrix}
	\bigscopta(x,y,z) =
		\begin{pmatrix}
	 		\scopa_0 \\ 0 \\ 0 \\ 0 \\ \vdots  \\
		\end{pmatrix},
}
where we note $\scopa_0 := \scopa_{0,0,0}(x,y,z) \equiv \genjac_0^{(a,0)} \: \ch_0$, and for each $n = 0,1,2\dots$,
\bseqn{
	A_n &:= 
		\begin{pmatrix}
			A^{(a)}_{x,n} \\
			A^{(a)}_{y,n} \\
			A^{(a)}_{z,n}
		\end{pmatrix} \in \R^{3(2n+1)\times(2n+3)}, \quad
	C_n := 
		\begin{pmatrix}
			C^{(a)}_{x,n} \\
			C^{(a)}_{y,n} \\
			C^{(a)}_{z,n}
		\end{pmatrix} \in \R^{3(2n+1)\times(2n-1)} \quad (n \ne 0), \nonumber \\
	B_n &:= 
		\begin{pmatrix}
			B^{(a)}_{x,n} \\
			B^{(a)}_{y,n} \\
			B^{(a)}_{z,n}
		\end{pmatrix} \in \R^{3(2n+1)\times(2n+1)}, \quad
	G_n(x,y,z) := 
		\begin{pmatrix}
			xI_{2n+1} \\
			yI_{2n+1} \\
			zI_{2n+1}
		\end{pmatrix} \in \R^{3(2n+1)\times(n+1)}.
}

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{2n+3}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence:
\bseqn{
	\begin{cases}
		\bigscopta_{-1}(x,y,z) := 0 \\
		\bigscopta_{0}(x,y,z) := \scopa_0 \\
		\bigscopta_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z), \quad n = 0,1,2,\dots.
	\end{cases}
}

@@

\begin{lemma}\label{lemma:sc:Rsecondderivative}
	For any general parameter $a$ and any $n = 0,1,\dots$, $k = 0,\dots,n$ we have that
\bseqn{
	&\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] \\
	&\quad \quad = \genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime \prime} - 2(k+1)z \genjacw^{(a+1, 2k)} \: \genjacnmk^{(a,2k) \: \prime} + (a+1) \genjacw^{(a, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime} \\
	&\quad \quad = \sum_{m = n-1}^{n+1} \: c_{m,k} \: \genjacw^{(a, 2k)} \: \genjac_{m-k}^{(a,2k)}
}
where 
\bseqn{
	c_{m,k} = - \frac{1}{\normgenjac^{(a,2k)}} \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z
}
\end{lemma}

@@

We now proceed with the case for the operator $D_\varphi^{(a)}$ for partial differentiation by $\varphi$. The entries of the operator are given by the coefficients in the expansion 
$$
\rhoppphi \scopnkia = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a+1)},
$$ 
where the coefficients are 
\bseqn{
	c_{m,j,h} = \norm{\scopmjh^{(a+1)}}^{-2}_{\scop^{(a+1)}} \ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}}.
}
Now, note that:
\bseqn{
	\genjacw^{(a,b) \: \prime}(z) &= a \: \genjacw^{(a-1,b)}(z) + c \: \rho(z) \: \rho'(z) \:\genjacw^{(a,b-2)}(z), \\
	\rho(z) \: \rho'(z) &= -z, \\
	\ppphi \scopnkia(x,y,z) &= -\rho(z) \: \ddz \Big[ \rho(z)^k \: \genjacnmk^{(a,2k)}(z) \Big] \chki(\theta), \\
	\ppphi \Big[ \genjacw^{(a,0)}(z) \: \scopnkia(x,y,z) \Big] &= -\rho(z) \: \ddz \Big[ \genjacw^{(a,k)}(z) \: \genjacnmk^{(a,2k)}(z) \Big] \chki(\theta).
}
Then, 
\bseqn{
	&\ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}} \\
	&\quad = - \int_\alpha^1 \Big( \int_0^{2\pi} \: \rho(z)^2 \ddz \: [\genjacnmk^{(a,2k)}(z) \: \rho(z)^k] \: \genjacmmj^{(a+1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \genjacw^{(a+1, 0)} \: \D z \\
	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \: \Big( \int_\alpha^1 \: \genjacmmj^{(a+1,2j)} \: [k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a+1, k + j)} \: \D z \Big) \\
	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(a+1,2k)} \: [k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime}] \: \genjacw^{(a+1, 2k)} \: \D z \\
	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} + \genjac_{m-k}^{(a+1,2k) \: \prime} \: \genjacw^{(1, 2)}\\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + a \: \rho^2 \: \genjac_{m-k}^{(a+1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
}
which is zero for $j \ne k$, $h \ne i$, and $m < n - 2$ by orthogonality.

>>>>

		& \quad + \alphaa_{n,k,3} \:  \scopa_{n, k-1, i}(x, y, z) + \alphaa_{n,k,4} \:  \scopa_{n, k+1, i}(x, y, z) \nonumber \\
		& \quad + \alphaa_{n,k,5} \:  \scopa_{n+1, k-1, i}(x, y, z) + \alphaa_{n,k,6} \:  \scopa_{n+1, k+1, i}(x, y, z), \\ \\

@@

		& \quad + \betaa_{n,k,i,3} \:  \scopa_{n, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,4} \:  \scopa_{n, k+1, |i-1|}(x, y, z) \nonumber \\
		& \quad + \betaa_{n,k,i,5} \:  \scopa_{n+1, k-1, |i-1|}(x, y, z) + \betaa_{n,k,i,6} \:  \scopa_{n+1, k+1, |i-1|}(x, y, z), \\ \\
	z \: \scopnkia(x,y,z) &= \gammaa_{n,k,1} \: \scopa_{n-1, k, i}(x, y, z) + \gammaa_{n,k,2} \: \scopa_{n, k, i}(x, y, z) \nonumber \\
		& \quad + \gammaa_{n,k,3} \: \scopa_{n+1, k, i}(x, y, z),

@@

Recall from \bsrefeqn{eqn:sc:normscop} that $\norm{\scopmjha}_{\scopa}^2 = \pi \: \normgenjac^{(a,2j)}$. Using a change of variables 
	(\cos\theta \sin\varphi, \: \sin\theta\sin\varphi, \: \cos\varphi) = (x, y, z)
we have that, for $m = 0,\dots,n+1$, $j = 0,\dots,m$, 

@@

	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \\
	&\quadfour \cdot \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \cos\theta \: \D \theta \Big) \\

@@

	&= \half \: \pi \: \delta_{i,h} \: (\eta_{k-1} \: \delta_{k-1, j} + \eta_{k} \: \delta_{k+1, j})  \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z

@@

	&= \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \rho(z)^{k+j+1} \: \genjacw^{(a, 0)}(z) \: \D z \Big) \\
	&\quadfour \cdot \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \sin\theta \: \D \theta \Big) \\

@@

	&= \half \: \pi \: \delta_{|i-1|,h} \: \big[(-1)^{i+1} \: \eta_{k-1} \: \delta_{k-1, j} + (-1)^i \: \eta_{k} \: \delta_{k+1, j}\big] \\
	&\quadfour \cdot \Big( \int^1_\alpha \genjacnmk^{(a, 2k)}(z) \: \genjacmmj^{(a, 2j)}(z) \: \genjacw^{(a, k + j + 1)}(z) \: \D z \Big)

@@

Each of these Jacobi matrices are then block-tridiagonal (block-bandwidths $(1,1)$). For $J_x^{(a)}$, the sub-blocks have sub-block-bandwidths $(2,2)$:
\bseqn{
	A^{(a)}_{x,n} &:= 
		\begin{pmatrix}
			0 & A^{(a)}_{n,0,6} & 0 & & \\
			A^{(a)}_{n,1,5} & \ddots & \ddots & & \\
			& \ddots & \ddots & \ddots & \\
			& & A^{(a)}_{n,n,5} & 0 & A^{(a)}_{n,n,6} \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)} \\
	B^{(a)}_{x,n} &:= 
		\begin{pmatrix}
			0 & A^{(a)}_{n,0,4} & & \\
			A^{(a)}_{n,1,3} & \ddots & \ddots & \\
			& \ddots & \ddots & A^{(a)}_{n,n-1,4} \\
			& & A^{(a)}_{n,n,3} & 0
		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)} \\
	C^{(a)}_{x,n} &:= 
		\begin{pmatrix}
			0 & A^{(a)}_{n,0,2} & & \\
			A^{(a)}_{n,1,1} & \ddots & \ddots & \\
			& \ddots & \ddots &A^{(a)}_{n,n-2,2} \\
			& & \ddots & 0 \\
			& & & A^{(a)}_{n,n,1} \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad (n \ne 0)
}
where for $k = 1,\dots,N, \: n = k,\dots,N$
\bseqn{
	A^{(a)}_{n,k,j} &:= 
		\begin{pmatrix}
			\alphaa_{n,k,j} & 0 \\
			0 & \alphaa_{n,k,j}
		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
	A^{(a)}_{n,0,j} &:=
		\begin{pmatrix}
			\alphaa_{n,0,j} & 0
		\end{pmatrix} \in \R^{1\times2}, j \text{ even} \\
	A^{(a)}_{n,1,j} &:=
		\begin{pmatrix}
			\alphaa_{n,1,j} \\
			0
		\end{pmatrix} \in \R^{2\times1}, j \text{ odd}.
}
For $J_y^{(a)}$, the sub-blocks have sub-block-bandwidths $(3,3)$:
\bseqn{
	A^{(a)}_{y,n} &:= 
		\begin{pmatrix}
			0 & B^{(a)}_{n,0,6} & 0 & & \\
			B^{(a)}_{n,1,5} & \ddots & \ddots & & \\
			& \ddots & \ddots & \ddots & \\
			& & B^{(a)}_{n,n,5} & 0 & B^{(a)}_{n,n,6} \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)} \\
	B^{(a)}_{y,n} &:= 
		\begin{pmatrix}
			0 & B^{(a)}_{n,0,4} & & \\
			B^{(a)}_{n,1,3} & \ddots & \ddots & \\
			& \ddots & \ddots & B^{(a)}_{n,n-1,4} \\
			& & B^{(a)}_{n,n,3} & 0
		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)} \\
	C^{(a)}_{y,n} &:= 
		\begin{pmatrix}
			0 & B^{(a)}_{n,0,2} & & \\
			B^{(a)}_{n,1,1} & \ddots & \ddots & \\
			& \ddots & \ddots & B^{(a)}_{n,n-2,2} \\
			& & \ddots & 0 \\
			& & & B^{(a)}_{n,n,1} \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad (n \ne 0)
}
where for $k = 1,\dots,N, \: n = k,\dots,N$  
\bseqn{
	B^{(a)}_{n,k,j} &:= 
		\begin{pmatrix}
			0 & \betaa_{n,k,0,j} \\
			\betaa_{n,k,1,j} & 0
		\end{pmatrix} \in \R^{2\times2}, (k \ne 1 \text { for } j \text{ odd}) \\
	B^{(a)}_{n,0,j} &:=
		\begin{pmatrix}
			0 & \betaa_{n,0,0,j}
		\end{pmatrix} \in \R^{1\times2}, j \text{ even} \\
	B^{(a)}_{n,1,j} &:=
		\begin{pmatrix}
			0 \\
			\betaa_{n,1,1,j}
		\end{pmatrix}\in \R^{2\times1}, j \text{ odd}.
}
For $J_z^{(a)}$, the sub-blocks are diagonal, i.e. have sub-block-bandwidths $(0,0)$:
\bseqn{
	A^{(a)}_{z,n} &:= 
		\begin{pmatrix}
			\Gamma^{(a)}_{n,0,3} & 0 & \\
			0 & \ddots & \ddots & & \\
			& \ddots & \ddots & \ddots & \\
			& & 0 & \Gamma^{(a)}_{n,n,3} & 0 \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n+3)} \\
	B^{(a)}_{z,n} &:= 
		\begin{pmatrix}
			\Gamma^{(a)}_{n,0,2} & \\
			& \ddots & & \\
			& & \ddots & \\
			& & & \Gamma^{(a)}_{n,n,2}
		\end{pmatrix} \in \R^{(2n+1)\times(2n+1)} \\
	C^{(a)}_{z,n} &:= 
		\begin{pmatrix}
			\Gamma^{(a)}_{n,0,1} & 0 & & \\
			0 & \ddots & \ddots & \\
			& \ddots & \ddots & 0 \\
			& & \ddots & \Gamma^{(a)}_{n,n-1,1} \\
			& & & 0 \\
		\end{pmatrix} \in \R^{(2n+1)\times(2n-1)}, \quad (n \ne 0)
}
where for $k = 1,\dots,N, \: n = k,\dots,N$
\bseqnnumber{
	\Gamma^{(a)}_{n,k,j} &:= 
		\begin{pmatrix}
			\gamma_{n,k,j} & 0 \\
			0 & \gamma_{n,k,j}
		\end{pmatrix} \in \R^{2\times2}, \label{eqn:sc:jacobisubblocksGamma1} \\
	\Gamma^{(a)}_{n,0,j} &:= \gammaa_{n,0,j}. \label{eqn:sc:jacobisubblocksGamma2}
}
Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs and the circular harmonics, meaning that the sparsity is not limited to the specific spherical cap, and would extend to the spherical band.

@@

A multivariate analogue of Clenshaw's algorithm for recursive evaluation of the OPs has been established for methods on other domains in 2D such as the triangle \cite{olver2019triangle} and the disk-slice \cite{snowball2019sparse}, and we proceed similarly here. Combining each system in \bsrefeqn{eqn:sc:jacobimatricesdefinitionalt} into a block-tridiagonal system, for any $(x,y,z) \in \Omega$, yields:
\bseqn{
	\renewcommand\arraystretch{1.3}
	&\begin{pmatrix}
		1 & & & \\
		B_0-G_0(x,y,z) & A_0 & & \\
		C_1 & B_1-G_1(x,y,z) & \quad A_1 \quad & \\
		& C_2 & B_2 - G_2(x,y,z)  & \ddots \\
		& & \ddots &\ddots
	\end{pmatrix}
	\bigscopta(x,y,z) \\
	&=
		\begin{pmatrix}
	 		\scopa_0 & 0 & 0 & 0 & \hdots
		\end{pmatrix}^\top,
}
where we note $\scopa_0 := \scopa_{0,0,0}(x,y,z) \equiv \genjac_0^{(a,0)} \: \ch_0$, and for each $n = 0,1,2\dots$,
\bseqn{
	A_n &:= 
		\begin{pmatrix}
			A^{(a)}_{x,n} \\
			A^{(a)}_{y,n} \\
			A^{(a)}_{z,n}
		\end{pmatrix} \in \R^{3(2n+1)\times(2n+3)}, \quad
	C_n := 
		\begin{pmatrix}
			C^{(a)}_{x,n} \\
			C^{(a)}_{y,n} \\
			C^{(a)}_{z,n}
		\end{pmatrix} \in \R^{3(2n+1)\times(2n-1)} \quad (n \ne 0), \nonumber \\
	B_n &:= 
		\begin{pmatrix}
			B^{(a)}_{x,n} \\
			B^{(a)}_{y,n} \\
			B^{(a)}_{z,n}
		\end{pmatrix} \in \R^{3(2n+1)\times(2n+1)}, \quad
	G_n(x,y,z) := 
		\begin{pmatrix}
			xI_{2n+1} \\
			yI_{2n+1} \\
			zI_{2n+1}
		\end{pmatrix} \in \R^{3(2n+1)\times(n+1)}.
}

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{2n+3}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
\bseqn{
	\begin{cases}
		\bigscopta_{-1}(x,y,z) := 0 \\
		\bigscopta_{0}(x,y,z) := \scopa_0 \\
		\bigscopta_{n+1}(x,y,z) = -\Dnt (B_n-G_n(x,y,z)) \bigscopta_n(x,y,z) - \Dnt C_n  \, \bigscopta_{n-1}(x,y,z)
	\end{cases}.
}

@@

\begin{lemma}\label{lemma:sc:Rsecondderivative}
	For any general parameter $a$ and any $n = 0,1,\dots$, $k = 0,\dots,n$ we have that
\bseqn{
	\ddx{z}[\genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime}] &= \genjacw^{(a+1, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime \prime} - 2(k+1)z \genjacw^{(a+1, 2k)} \: \genjacnmk^{(a,2k) \: \prime} \\
	&\quadfour + (a+1) \genjacw^{(a, 2(k+1))} \: \genjacnmk^{(a,2k) \: \prime} \\
	&= \sum_{m = n-1}^{n+1} \: c_{m,k} \: \genjacw^{(a, 2k)} \: \genjac_{m-k}^{(a,2k)}
}
where 
\bseqn{
	c_{m,k} = - \frac{1}{\normgenjac^{(a,2k)}} \: \int_\alpha^1 \: \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \: \D z
}
\end{lemma}

@@

We now proceed with the case for the operator $D_\varphi^{(a)}$ for partial differentiation by $\varphi$. The entries of the operator are given by the coefficients in the expansion 
$$
\rhoppphi \scopnkia = \sum_{m=0}^{n+1} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(a+1)},
$$ 
where the coefficients are 
\bseqn{
	c_{m,j,h} = \norm{\scopmjh^{(a+1)}}^{-2}_{\scop^{(a+1)}} \ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}}.
}
Now, note that:
\bseqn{
	\genjacw^{(a,b) \: \prime}(z) &= a \: \genjacw^{(a-1,b)}(z) + c \: \rho(z) \: \rho'(z) \:\genjacw^{(a,b-2)}(z), \\
	\rho(z) \: \rho'(z) &= -z, \\
	\ppphi \scopnkia(x,y,z) &= -\rho(z) \: \ddz \Big[ \rho(z)^k \: \genjacnmk^{(a,2k)}(z) \Big] \chki(\theta), \\
	\ppphi \Big[ \genjacw^{(a,0)}(z) \: \scopnkia(x,y,z) \Big] &= -\rho(z) \: \ddz \Big[ \genjacw^{(a,k)}(z) \: \genjacnmk^{(a,2k)}(z) \Big] \chki(\theta).
}
Then, 
\bseqn{
	&\ip<\rhoppphi \scopnkia, \: \scopmjh^{(a+1)}>_{\scop^{(a+1)}} \\
	&= - \int_\alpha^1 \Big\{ \int_0^{2\pi} \rho(z)^2 \ddz \:  \big[ \genjacnmk^{(a,2k)}(z) \: \rho(z)^k \big] \: \genjacmmj^{(a+1,2j)}(z) \: \rho(z)^j \: \chki(\theta) \: \chjh(\theta) \: \D \theta \\
	&\quadfour \quadtwo \cdot \genjacw^{(a+1, 0)}(z) \Big\} \: \D z \\
	&= \Big( \int_0^{2\pi} \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
	&\quadfour \cdot \Big( \int_\alpha^1 \: \genjacmmj^{(a+1,2j)} \: \big[ k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime} \big] \: \genjacw^{(a+1, k + j)} \: \D z \Big) \\
	&= \pi \delta_{k,j} \delta_{i,h} \int_\alpha^1 \genjac_{m-k}^{(a+1,2k)} \: \big[ k z \genjacnmk^{(a, 2k)} - \rho^2 \genjacnmk^{(a, 2k) \: \prime} \big] \: \genjacw^{(a+1, 2k)} \: \D z \\
	&= \pi \delta_{k,j} \delta_{i,h} \int_\alpha^1 \genjacnmk^{(a, 2k)} \: \Big\{  k z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} + \genjac_{m-k}^{(a+1,2k) \: \prime} \: \genjacw^{(1, 2)}\\
	&\quadeight + a \: \rho^2 \: \genjac_{m-k}^{(a+1,2k)} - (2k+2) z \: \genjac_{m-k}^{(a+1,2k)} \: \genjacw^{(1, 0)} \Big\} \: \genjacw^{(a, 2k)} \: \D z
}
which is zero for $j \ne k$, $h \ne i$, and $m < n - 2$ by orthogonality.
